YouTube Kids: The Effect of YouTube on Children’s Racial Learning

### 2025-02-27 manuscript.qmd
Updated manuscript draft

---
title: "YouTube Kids: The Effect of YouTube on Children’s Racial Learning"
shorttitle: "YouTube Racial Attitudes"
author: 
- name: Reilly Amera
  corresponding: true
  email: ramera@uchicago.edu
  affiliations:
  - id: id1
    name: "University of Chicago"
    department: MA Program in the Social Sciences
bibliography: [r-references, bibliography.bib]
format: pdf
editor: visual
---

@@

library(papaja)
library(knitr)
library(broom)
library(tidyverse)

# Generate or overwrite r-reference files
r-refs("r-references.bib")
```

@@

Not only does ethnic-racial socialization occur with explicit instruction, but so too do children attend to the racial patterns and organizations embedded into their environments and social settings [@hugheswatford_2022_racial]. Such environmental racial patterns—who tends to hold positions of authority, who tends to receive discipline in school, etc.—constitute racial regularities, or “persistent patterns in the relative social experiences of, or roles occupied by, two or more racial (or racialized groups) within a setting”  [@hugheswatford_2022_racial]. In response to such patterns, children engage in racial learning, or the meaning-making of external messaging so as to form their own conceptions of how race works in their social world [@hugheswatford_2022_racial]. As such, when exposed to consistent, repeated regularities, children draw inferences as to why such patterns exist and persist; they may accept that such patterns as necessary features of ‘the way things are,’ or may even create explanations for such patterns (e.g. “White families live in big houses because they work hard” in contrast to “Black families live in big houses because they got lucky”) [@hughesetal_2023_growing; @rizzoetal_2022_beliefs]. Further, this learning may be yet more implicit: when tested in the same environment, young children show levels of implicit biases, unconscious mental associations based on race or other social categories, similar to those of adults [@payneetal_2017_bias]. This suggests that features of specific environments such as racial regularities not only affect explicitly articulated beliefs but implicit understandings of one’s social environment. As such, to fully understand how children’s conceptions of race form, we must examine the ways in which environments containing ontologically inaccurate racial regularities affect their racial learning.

@@

```{r}
#| label: data analysis
#| echo: false
# FOR THIS DRAFT - only include child data and simple regressions 
```

# References
::: {#refs}
:::

>>>>

---
title: "YouTube Kids: The Effect of YouTube on Children’s Racial Learning"
shorttitle: "YouTube Racial Attitudes"
author: 
- name: Reilly Amera
  corresponding: true
  email: ramera@uchicago.edu
  affiliations:
  - id: id1
    name: "University of Chicago"
    department: MA Program in the Social Sciences
bibliography: [bibliography.bib]
link-citations: true
draft-date: false
lang: en
language:
  citation-last-author-separator: "and"
  citation-masked-author: "Masked Citation"
  citation-masked-date: "n.d."
  citation-masked-title: "Masked Title"
  email: "Email"
format:
  apaquarto-html: default
  apaquarto-docx: default
  apaquarto-typst: 
    keep-typ: true
    list-of-figures: true
    list-of-tables: true
    toc: true
  apaquarto-pdf:
    documentmode: man
    keep-tex: true
    list-of-figures: true
    list-of-tables: true
    fig-width: 2
    fig-height: 2
    out-width: 40%
editor: visual
---

@@

library(svglite)
library(papaja)
library(knitr)
library(broom)
library(tidyverse)

# Generate or overwrite r-reference files
# papaja::r_refs(file = "r-references.bib")

```

@@

Not only does ethnic-racial socialization occur with explicit instruction, but so too do children attend to the racial patterns and organizations embedded into their environments and social settings [@hugheswatford_2022_racial]. Such environmental racial patterns—who tends to hold positions of authority, who tends to receive discipline in school, etc.—constitute racial regularities, or “persistent patterns in the relative social experiences of, or roles occupied by, two or more racial (or racialized groups) within a setting” [@hugheswatford_2022_racial]. In response to such patterns, children engage in racial learning, or the meaning-making of external messaging so as to form their own conceptions of how race works in their social world [@hugheswatford_2022_racial]. As such, when exposed to consistent, repeated regularities, children draw inferences as to why such patterns exist and persist; they may accept that such patterns as necessary features of ‘the way things are,’ or may even create explanations for such patterns (e.g. “White families live in big houses because they work hard” in contrast to “Black families live in big houses because they got lucky”) [@hughesetal_2023_growing; @rizzoetal_2022_beliefs]. Further, this learning may be yet more implicit: when tested in the same environment, young children show levels of implicit biases, unconscious mental associations based on race or other social categories, similar to those of adults [@payneetal_2017_bias] This suggests that features of specific environments such as racial regularities not only affect explicitly articulated beliefs but implicit understandings of one’s social environment. As such, to fully understand how children’s conceptions of race form, we must examine the ways in which environments containing ontologically inaccurate racial regularities affect their racial learning. 

@@

## Methods {#sec-methods}

### Participants and Procedures {#sec-participants-and-procedures}
This study will include children between the ages of 8 and 13. Though children may express negative racial attitudes prior to this point in development, during early adolescence, children develop the ability to recognize and reflect on the racial patterns and dynamics present in their environments (Hughes & Watford, 2022). As such, this age range should allow me to observe how the observation racial regularities present on YouTube may affect children’s stereotyping and prejudicial attitudes, and how such a phenomenon develops. Participants will be recruited through Qualtrics Panels, which will allow me to obtain a representative sample of parents and children; other studies have used Qualtrics Panels with similar methods and populations (Byrd & Ahn, 2020). 
Data will be collected during February and March of 2025 from families recruited via Qualtrics Panels. Participating families will receive two Qualtrics surveys to be completed asynchronously, and they will receive compensation (around $10-$15; the final amount will be determined by Qualtrics Panels) after both surveys have been completed. First, parents will complete a demographics and YouTube usage survey, in which they will answer questions related to parent and child demographics (child age, race/ethnicity, parent income and level of education, etc), as well as questions regarding their child’s YouTube usage (for instance, time spent on YouTube, as reported by their child’s device screen time). 
Then, children will complete their own survey.

### Measures {#sec-measures}
First, they will answer questions related to their YouTube usage (perceptions of time spent). Next, they will either answer more questions pertaining to YouTube content, or their general racial attitudes; either block of questions will appear randomly, so as to reduce the influence of question order. In the racial attitudes section, children will answer questions regarding outgroup prejudices, as rated on a Likert scale. They will also complete one friend choice item, as a simple exploratory measure of racial preference. In the YouTube content section, children will be asked to think of their favorite YouTuber as they answer questions related to whether they have a parasocial relationship with that YouTuber, as well as their observations of racial regularities on YouTube.

## Results {#sec-results}

```{r}
#| label: regression 
#| echo: false
#| message: false
#| warning: false

# In the future, will have a final, written out, intermediate dataset that will be read in here.

# regression model
model <- lm(malleable_score ~ Ptime, data = youtube)

```

```{r}
#| label: regression-report-functions
#| echo: false
#| message: false
#| warning: false

# report_regression takes a lm model object as an argument
# and returns a string of format:
# "$F$(df, df.residual) = F-value, $p$ = p-value, $R^2$ = r-squared"
report_regression <- function(model) {

  # Extract coefficients and p-values
  m_glance <- glance(model)

  # Format the p-value
  p_value <- apa_p(m_glance$p.value, add_equals = TRUE)

  # Create a string with the results
  report <- paste0("$F$(", 
                   m_glance$df, 
                   ", ", m_glance$df.residual, 
                   ") = ", round(m_glance$statistic, 2), 
                   ", $p$ = ", p_value, ", $R^2$ = ", round(m_glance$r.squared,2))

  return(report)

}

```

Below is a simple regression displaying the relationship between time spent on YouTube and a child's racial attitudes. Despite a visual upward trend, this relationship is not statistically significant: `r report_regression(model)`.

```{r}
#| label: regression-plot
#| fig-cap: "Simple Correlation Between Time Spent on YouTube and Racial Attitudes score"
#| fig-scap: "Time v. Attitudes"
#| fig-align: "center"
#| fig-width: 6
#| fig-height: 4
#| echo: false
#| message: false
#| warning: false

model_plot <- ggplot(youtube) +
  aes(x = Ptime, y = malleable_score) +
  geom_point() +
  stat_smooth(method = "lm") +
  scale_y_continuous(name = "Racial Attitudes Score") +
  scale_x_continuous(name = "Time spent on YouTube")
model_plot

```

# References
